---
title: "US Beers and Breweries"
author: "Christopher Morgan, Dmitry Andreev, & Eduardo Cantu"
date: "February 17, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
--Intro--

## Data Description

### Data Load
```{r, Data Load}
#Read Data into Variables
beers <- read.csv("Beers.csv")
breweries <-read.csv("Breweries.csv")

#Rename beers name column to beerName
colnames(beers)[1] <- c("beerName")

#Rename brewery name column to brewerCompany
colnames(breweries)[2] <- c("brewerCompany")

#Rename Brewery_id to Brew_id in beers data, same as in breweries in order to merge data sets
colnames(beers)[5] <- c("Brew_ID")

# Trim the leading white-spce from the State
breweries$State <- trimws(breweries$State)

#Create State Names data frame, inlcuding Washington DC
stateNames <- data.frame(stateName=c(state.name,"Washington DC"),State=c(state.abb,"DC"))


# add the State Name to the Breweries Data Frame
breweries <- merge(breweries,stateNames,by="State", all=TRUE)


```
# Questions of Interest

##1. How many breweries are present in each state?
```{r, CountBreweries}
# Count Breweries per State

countBrewByState <- table(breweries$stateName)
countBrewByState <- as.data.frame(countBrewByState)
colnames(countBrewByState)=c("stateName","BreweriesCount")
countBrewByState
```

##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r, Merge Data, cols.print=11}
#Merge beers and breweries data together
m1 <- merge(beers,breweries,by="Brew_ID")

#Table with the first 6 observations from m1 (merged data)
head(m1)

#Talble with the last 6 observations from m1 (merged data)
tail(m1)

```

##3. Report the number of NA's in each column.

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

##6. Summary statistics for the ABV variable.

##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship heand EXPLAIN your answer.

#Conclusion
--Conclusion--

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
t