---
title: "BeerVana Consulting"
author: "Christopher Morgan, Dmitry Andreev, & Eduardo Cantu"
subtitle: "US Beers and Breweries"
date: "February 28, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
As we enter 2018, with nearly 6,000 microbreweries registered in the U.S., it is safe to say that craft brews are all the craze and they're here to stay. When was the last you can remember reaching for a Miller Light after leaving college? Sure, this could be due to the fact you can finally afford to pay more than $1.50 for a canned beer and are starting to get curious about trying something new. At BeerVana Consulting we believe that the real driving force behind this trend is the unmatched variety of flavors that you simply can't find with the Budweisers, Coors, and Heinekens of the world. Not only are there an endless array of hole-in-the-wall breweries popping up on every town corner but also the ever-increasing competition to make new brews that cater to this new class of beer snob elites has never been fiercer. Lucky for them, we have narrowed down the main preferences for each state. So for any new players looking to enter the craft brewery game, rest assured that Beervana is taking out the guesswork for you and your next brew! The goal of this report is to simplify the large sea of information currently available for some of the most popular craft breweries in each state and the most popular canned beers they're putting out. We also dive into a rather surprising relationship that exists between the ABV (Alcohol by Volume) and IBU (International Bitterness Unit) for these beers.

## Data Description
The datasets being used for this report was sourced from the CraftCans website. It only contains data for canned beers from craft breweries in the United States. It is not clear from the website if this dataset reports every single canned beer brewed in the US or not. There are also several attributes of the listed beers that have missing information so for this reason we will consider this dataset to be a sample that may contain biases.

The two datasets that will be referred and used throughout this report are summarized below:

1. Beers.csv
    + Name: Name of the beer.
    + Beer ID: Unique identifier of the beer.
    + ABV: Alcohol by volume of the beer.
    + IBU: International Bitterness Units of the beer.
    + Brewery ID: Brewery id associated with the beer.
    + Style: Style of the beer.
    + Ounces: Ounces of beer.

2. Breweries.csv
    + Brew ID: Unique identifier of the brewery.
    + Name: Name of the brewery.
    + City: City where the brewery is located.
    + State: U.S. State where the brewery is located.

### Data Load
```{r, Data Load}
#Read Data into Variables
beers <- read.csv("Beers.csv")
breweries <-read.csv("Breweries.csv")

#Rename beers name column to beerName
colnames(beers)[1] <- c("beerName")

#Rename brewery name column to brewerCompany
colnames(breweries)[2] <- c("brewerCompany")

#Rename Brewery_id to Brew_id in beers data, same as in breweries in order to merge data sets
colnames(beers)[5] <- c("Brew_ID")

# Trim the leading white-spce from the State
breweries$State <- trimws(breweries$State)

#Create State Names data frame, inlcuding Washington DC
stateNames <- data.frame(stateName=c(state.name,"Washington DC"),State=c(state.abb,"DC"))


# add the State Name to the Breweries Data Frame
breweries <- merge(breweries,stateNames,by="State", all=TRUE)


```
# Questions of Interest

##1. How many breweries are present in each state?
```{r, CountBreweries}
# Count Breweries per State

countBrewByState <- table(breweries$stateName)
countBrewByState <- as.data.frame(countBrewByState)
colnames(countBrewByState)=c("stateName","BreweriesCount")
countBrewByState
```

##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r, Merge Data, cols.print=11}
#Merge beers and breweries data together
m1 <- merge(beers,breweries,by="Brew_ID")

#Table with the first 6 observations from m1 (merged data)
head(m1)

#Talble with the last 6 observations from m1 (merged data)
tail(m1)

```

##3. Report the number of NA's in each column.

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

##6. Summary statistics for the ABV variable.

##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship heand EXPLAIN your answer.

#Conclusion
--Conclusion--

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
t