---
title: "US Beers and Breweries"
author: "Christopher Morgan, Dmitry Andreev, & Eduardo Cantu"
date: "February 17, 2018"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
--Intro--

## Data Description
The Beers dataset contains a list of 2410 US craft beers and Breweries dataset contains 558 US
breweries. The datasets descriptions are as follows.


### Beers.csv:

Field Name | Field Description
-----------|--------------------
Name:      | Name of the beer.
Beer ID:   | Unique identier of the beer.
ABV:       | Alcohol by volume of the beer.
IBU:       | International Bitterness Units of the beer.
Brewery ID:| Brewery id associated with the beer.
Style:     | Style of the beer.
Ounces:    | Ounces of beer.


### Breweries.csv:
Field Name | Field Description
-----------|--------------------
Brew ID:   | Unique identier of the brewery.
Name:      | Name of the brewery.
City:      | City where the brewery is located.
State:     | U.S. State where the brewery is located.

### Data Load
```{r, Data Load}
#Read Data into Variables
beers <- read.csv("Beers.csv")
breweries <-read.csv("Breweries.csv")

#Rename beers name column to beerName
colnames(beers)[1] <- c("beerName")

#Rename brewery name column to brewerCompany
colnames(breweries)[2] <- c("brewerCompany")

#Rename Brewery_id to Brew_id in beers data, same as in breweries in order to merge data sets
colnames(beers)[5] <- c("Brew_ID")

# Trim the leading white-spce from the State
breweries$State <- trimws(breweries$State)

#Create State Names data frame, inlcuding Washington DC
stateNames <- data.frame(stateName=c(state.name,"Washington DC"),State=c(state.abb,"DC"))

# add the State Name to the Breweries Data Frame
breweries <- merge(breweries,stateNames,by="State", all=TRUE)

```
# Questions of Interest

##1. How many breweries are present in each state?
```{r, CountBreweries}
# Count Breweries per State

countBrewByState <- table(breweries$stateName)
countBrewByState <- as.data.frame(countBrewByState)
colnames(countBrewByState)=c("stateName","BreweriesCount")
countBrewByState
```

##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

```{r, Merge Data, cols.print=11}
#Merge beers and breweries data together
m1 <- merge(beers,breweries,by="Brew_ID")
```

with the merged data, we can begin a quick pass at removing duplicates in our dataset

```{r, Clean Merged Data, cols.print=11}

#performs search for parenthesis values, removes them from a nondestructive column for beers
m1$beerNameNoY<-gsub("[:(:].*[:):]","",m1$beerName)
#trims white space around newly trimmed data for beers
m1$beerNameNoY<-trimws(m1$beerNameNoY)
#performs search for parenthesis values, removes them from a nondestructive column for brewries
m1$brewerNameNoP<-gsub("[:(:].*[:):]","",m1$brewerCompany)
#trims white space around newly trimmed data for brewries
m1$brewerNameNoP<-trimws(m1$brewerNameNoP)

#eliminates duplicates for beers
m1 <- m1[!duplicated(m1[,c("beerNameNoY","brewerNameNoP","ABV","IBU","Style")]),]

#Table with the first 6 observations from m1 (merged data)
head(m1)

#Talble with the last 6 observations from m1 (merged data)
tail(m1)

```


##3. Report the number of NA's in each column.
``` {r, NACountCalculation}
# This will count NA per field
NACount <- sapply(m1,function(x) sum(is.na(x)))
NACount <- as.data.frame(NACount)
colnames(NACount)=c("CountOfNAs")
NACount


```

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r, medianCalculation}
# ABV Median Calculation per State and Column Renames. NA Values are ignored.
med.st.abv<-aggregate(m1$ABV~m1$State,m1,median,na.rm=T)
colnames(med.st.abv)<- c("State","ABV")

# IBU Median Calculation per State and Column Renames. NA Values are ignored.
med.st.ibu<-aggregate(m1$IBU~m1$State,m1,median,na.rm=T)
colnames(med.st.ibu)<- c("State","IBU")

# ABV Median per State 
med.st.abv

# IBU Median per State
med.st.ibu

library(ggplot2)

##Histograms, Raw Data

ggplot(data=med.st.abv, aes(x=med.st.abv$State,y=med.st.abv$ABV)) + geom_bar(stat="identity")
ggplot(data=med.st.ibu, aes(x=med.st.ibu$State,y=med.st.ibu$IBU)) + geom_bar(stat="identity")

####NOTE SD DOES NOT SHOW UP IN IBU BECAUSE OF MISSING DATA, WHAT DO WE DO WITH THIS???


```


##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r MaxABV}
# Find the State with the maximum ABV beer
z<-which.max(m1$ABV)

paste(m1$stateName[z], "has the highest ABV with a Max ABV of:", m1$ABV[z])

# Find the State with the highest bitterness (IBU) beer
y<-which.max(m1$IBU)
paste(m1$stateName[y], "has the highest IBU with a Max IBU of:", m1$IBU[y])


```

##6. Summary statistics for the ABV variable.
``` {r, ABVSummary}
# Summary for ABV Values
summary(m1$ABV)
```
##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship heand EXPLAIN your answer.
``` {r Relationship}

# Scatter plot for bitternes vs alcohol content
ggplot(data=m1,aes(x=m1$ABV,y=m1$IBU)) + geom_point() + geom_smooth(method="lm") + xlab("abv") + ylab("Bitternes") + ggtitle("Beer Bitternes vs Alcohol Content")

```

#Conclusion

--Conclusion--

