---
title: "US Beers and Breweries"
author: "Christopher Morgan, Dmitry Andreev, & Eduardo Cantu"
date: "February 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
--Intro--

## Data Description

### Data Load
```{r, Data Load}
#Read Data into Variables
beers <- read.csv("Beers.csv")
breweries <-read.csv("Breweries.csv")

```
# Questions of Interest

##1. How many breweries are present in each state?

##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

##3. Report the number of NA's in each column.

##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

##6. Summary statistics for the ABV variable.

##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot. You are welcome to use the ggplot2 library for graphs. Please ignore missing values in your analysis. Make your best judgment of a relationship and EXPLAIN your answer.

#Conclusion
--Conclusion--

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
